Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

\title*{Challenges in Developing Great Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al}
\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and Aleksei G. Sorokin}
\institute{Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and Yuhan Ding \at
\and
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

@@

\newpage

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ a low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean
    \begin{equation}
        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten fit above form and/or to facilitate a good approximation with small $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a desired error tolerance.

@@

\begin{itemize}
\item Integrated with related libraries (Sect.\ \ref{CDHS_sec:integrated}),

\item Correct (Sect.\ \ref{CDHS_sec:correct}),

\item Efficient in computational time and memory (Sect.\ \ref{CDHS_sec:efficient}),

\item Accessible to practitioners and theorists alike (Sect.\ \ref{CDHS_sec:accessible}),

\item Sustainable by a community that owns it (Sect.\ \ref{CDHS_sec:sustainable}), and

\item Scalable to advanced hardware architectures (Sect.\  \ref{CDHS_sec:scalable}).

@@

Developing software to implement such enhanced MC strategies poses a number of challenges. First, a development team must determine the value of the software to the users. We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHS_sec:why_we_need_software}. Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHS_sec:tenets_of_great_software}. Third, an architecture for the  (Q)MC software should be determined. We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHS_sec:qmc_software_architecture}. Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support. We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHS_sec:language_library_environment}. Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community. To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.

\section{Why Develop QMC Software} \label{CDHS_sec:why_we_need_software}

@@

QMC software makes theoretical advances in QMC available to practitioners. However, translating pseudo-code into good executable code is nuanced.  Software developers must write code that is computationally efficient, numerical stable, and provides reasonable default choices of tuning parameters. Good QMC software relieves users of these concerns.

@@

Good QMC software opens up new application areas for QMC methods by allowing practitioners to compare new methods to existing ones.  QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering. Unexpectedly good or bad computational results leads to open theoretical questions.  For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.

@@

\section{Integrated} \label{CDHS_sec:integrated}

Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}$:
\begin{gather}
\label{eq:cubSummary}
	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsx_i) =: \hat{\mu}_n, \\
 \label{CDHS_eq:error_tol}
 \abs{\mu - \hat{\mu}_n} \le \varepsilon.
\end{gather}
Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.  Great QMC software environments allows different implementations of these components to be interchanged.

@@

For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, the LD sequence generators included already allow users to easily access them while drawing from other packages in these environments.  Some large environments, like Dakota \cite{DakotaUsersManual} and SAS have limited LD generators and should expand their offerings.  Developers of smaller libraries, like BRODA and QMCPy, should take initiative to show how their libraries can connect well with others.

\subsection{Integrands and Variable Transformations} The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
\begin{itemize}
	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ is the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},

	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$ \cite{}, or

	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest.
\end{itemize}

To put $\mu$ into the form that is directly accessible to QMC methods requires a \emph{variable transformation}, $\bsPsi$, such that
\begin{equation*}
    \Psi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
\end{equation*}
Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.  If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.  Moreover, one might choose $\bsPsi$ such that $\bsT = \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.

@@

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UM-Bridge \cite{umbridge} (the UQ and Model Bridge), which connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.  We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a} where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.  The beam position is discretized at $30$ points.  QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of the laptop computer to decrease execution time.

@@

\subsection{Stopping Criteria}
The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.  See \cite{HicJim16a} for details.  Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.  Some of these stopping criteria for absolute error criteria, \eqref{CDHS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.

@@

\section{Correct} \label{CDHS_sec:correct}
Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.  Here are two examples.

@@

\section{Efficient} \label{CDHS_sec:efficient}
Another obvious feature of great QMC software is numerical efficiency.  This requires developers to identify bottlenecks in performance.  Intensive computation may need to be rewritten in lower level languages, like C.  Code may need to be refactored to avoid excessive calls to memory.  As new hardware architectures arise, QMC software should adapt.

\section{Accessible} \label{CDHS_sec:accessible}
One attribute of great QMC software that often receives insufficient attention is its accessibility.  Developers may write code for their own purposes and move on.  Great software should be stored on a repository, where the latest version can be downloaded.  There should be documentation that guides the user through installation and explains the library's features.  Demos are crucial for showing how the software work and forming a basis for a new user to write code to solve their own examples.  There should be a place for users to report bugs and make feature requests.  Tutorials help newcomers to QMC understand its advantages. Blogs educate a wider audience and enlarge our QMC community.

\section{Sustainable} \label{CDHS_sec:sustainable} 
Great QMC software should be alive.  It should be under active development or merged into a libraary that is actively maintained.  Commercial software has a infrastructure to sustain the algorithms in their libraries, but without attention, those algorithms can become stale.  Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.

@@

\section{Tenets of Great Software} \label{CDHS_sec:tenets_of_great_software}

Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication. A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement. Software accessibility considerations are further discussed in Sect.\ \ref{CDHS_sec:language_library_environment}.

Libraries should be \emph{sustainable} for future development. The goal should be to have a sufficient user base and developer community for updates and maintenance. Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section. However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support. We promote collaborative, community driven development further in Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.

\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHS_sec:qmc_software_architecture}

\section{Choosing a Environment} \label{CDHS_sec:language_library_environment}

\section{Encouraging Collaborative Development} \label{CDHS_sec:encouraging_collaborative_developement}

\section{Example Collaboration with UMBridge and QMCPy} \label{CDHS_sec:example_collaboration_umbridge}

@@

\AGSComment{
\tableofcontents
\section*{TODO}
\begin{itemize}
    \item Flesh out Section \ref{CDHS_sec:encouraging_collaborative_developement}
    \item QMC Software Architecture section: better way then description of components and list of softwares?
    \item For Section \ref{CDHS_sec:tenets_of_great_software}, the slides also include
    \begin{itemize}
        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
        \item Current---include the latest and best algorithms
    \end{itemize}
    which I think fall under the umbrella of software being \emph{integrated} with the community.
    \item update QMCPy citation \cite{QMCPy2020a}
    \item figures?
    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
\end{itemize}
}

@@

\section{Introduction} \label{CDHS_sec:intro}
Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.  In some cases QMC  achieves one hundredth of the error of IID MC in the same amount of time. Often, these efficiency gains are obtained simply by  replacing IID sampling by low discrepancy (LD) sampling, which is the heart of QMC.

@@

\section{Available Software for QMC} \label{CDHS_sec:available}
QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
Software focusing on generating high quality LD sequences  or their generators includes
\begin{description}[format=\textup,format=\textbf]
% https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
	\item[BRODA] Sobol' sequences in C, MATLAB, and Excel \cite{BRODA20a},
	\item[Burkhardt] various QMC software in C++, Fortran, MATLAB, \& Python \cite{Bur20a},
	\item[LatNet Builder] Generating vectors/matrices for lattices and digital nets \cite{LEcEtal22a,LatNet},
	\item[MATLAB] Sobol' and Halton sequences, commercial \cite{MAT9.9},
	\item[MPS] Magic Point Shop, lattices and Sobol' sequences \cite{Nuy17a},
	\item[Owen] Randomized Halton sequences in R \cite{Owe20a},
	\item[PyTorch] Scrambled Sobol' sequences \cite{PyTorch},
	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}, and
	\item [qrng]  Sobol', Halton, and Korobov sequences in R \cite{QRNG2020}.
\end{description}
Software focusing on QMC algorithms and applications includes
\begin{description}[format=\textup,format=\textbf]
	\item[GAIL] Automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal20a},
	\item[ML(Q)MC] Multi-Level (Quasi-)Monte Carlo routines in C++, MATLAB, Python, and R \cite{GilesSoft},
	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks 'N Statistics in Python \cite{OpenTURNS},
	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients \cite{KuoNuy16a},
	\item[SSJ] Stochastic Simulation in Java \cite{SSJ}, and
	\item[UQLab] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.
\end{description}

@@

The software libraries referenced in Section \ref{CDHS_sec:available} provide one or more of these components. QMCPy combines multiple examples of them all under and object oriented framework. Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component. The following sections detail descriptions and use cases for each component and select implementations.

@@

\section{Stopping Criteria} \label{CDHS_sec:stopping_crit}

@@

QMC, which uses LD sequences, is touted as providing substantially greater computational efficiency compared to IID MC.
Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences. Consistent with what is stated in Section \ref{CDHS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.  By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.  We see that QMC methods often require orders of magnitude fewer samples that MC methods to achieve the same error tolerance.

@@

A helpful features of QMCPy is the ability to print out components instances to see more detailed information. This allows users to quickly see the properties that have been set and defaulted for any given component. This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHS_sec:stopping_crit}. The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}

@@

\section{Further Work} \label{CDHS_sec:further}

@@

>>>>

\title*{Challenges in Developing Great Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al}
\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and R. Jagadeeswaran Aleksei G. Sorokin}
\institute{
Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
\and 
Yuhan Ding \at
 Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{yding2@hawk.iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and
R. Jagadeeswaran \at ???
\and
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

@@

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ a low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean
    \begin{equation}
        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit above form and/or to facilitate a good approximation with small $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

\begin{itemize}
\item Integrated with related libraries (Sect.\ \ref{CDHJS_sec:integrated}),

\item Correct (Sect.\ \ref{CDHJS_sec:correct}),

\item Efficient in computational time and memory (Sect.\ \ref{CDHJS_sec:efficient}),

\item Accessible to practitioners and theorists alike (Sect.\ \ref{CDHJS_sec:accessible}), and

\item Sustainable by a community that owns it (Sect.\ \ref{CDHJS_sec:sustainable}).

@@

In each section we describe the challenges faced and how they might be overcome.

\section{Why Develop QMC Software} \label{CDHJS_sec:why_we_need_software}

@@

QMC software makes theoretical advances in QMC available to practitioners. However, translating pseudo-code into good executable code is nuanced.  Software developers must write code that is computationally efficient, numerical stable, and provides reasonable default choices of tuning parameters. Great QMC software relieves users of these concerns.

@@

Great QMC software opens up new application areas for QMC methods by allowing practitioners to compare new methods to existing ones.  QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering. Unexpectedly good or bad computational results leads to open theoretical questions.  For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.

@@

\section{Integrated} \label{CDHJS_sec:integrated}

Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}$:
\begin{gather}
\label{eq:cubSummary}
	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsx_i) =: \hat{\mu}_n, \\
 \label{CDHJS_eq:error_tol}
 \abs{\mu - \hat{\mu}_n} \le \varepsilon.
\end{gather}
Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.  Great QMC software environments allows different implementations of these components to be freely interchanged.

@@

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these could provide more flexibility in their LD offerings.  Some large environments, such as  Dakota \cite{DakotaUsersManual} and SAS, have quite limited LD generators and should expand their offerings.  Developers of smaller libraries with more QMC features, like BRODA and QMCPy, should demonstrate how their libraries can connect well with other software environments.  An example involving QMCPy is given in the next subsection.

\subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
\begin{itemize}
	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ is the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},

	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$ \cite{KuoNuy16a}, or

	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest \cite{GelEtal13}.
\end{itemize}

To put $\mu$ into the form that is directly accessible to QMC methods requires a variable transformation, $\bsPsi$, such that
\begin{equation*}
    \Psi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
\end{equation*}
Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.  If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.  Moreover, one might choose $\bsPsi$ such that $\bsT = \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.

@@

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.  We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.  The beam position is discretized at $30$ points.  QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of the laptop computer to decrease execution time.

@@

\subsection{Stopping Criteria}
The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.  See \cite{HicJim16a} for details.  Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.  Some of these stopping criteria for absolute error criteria, \eqref{CDHJS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.

@@

\section{Correct} \label{CDHJS_sec:correct}
Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.  Here are two examples.

@@

\section{Efficient} \label{CDHJS_sec:efficient}
Another obvious feature of great QMC software is numerical efficiency.  This requires developers to identify bottlenecks in performance.  Intensive computation may need to be rewritten in lower level languages, like C.  Code may need to be refactored to avoid excessive calls to memory.  As new hardware architectures arise, QMC software should adapt.

\section{Accessible} \label{CDHJS_sec:accessible}
One attribute of great QMC software that often receives insufficient attention is its accessibility.  Developers may write code for their own purposes and move on.  Great software should be stored on a repository, where the latest version can be downloaded.  There should be documentation that guides the user through installation and explains the library's features.  Demos are crucial for showing how the software work and forming a basis for a new user to write code to solve their own examples.  There should be a place for users to report bugs and make feature requests.  Tutorials help newcomers to QMC understand its advantages. Blogs educate a wider audience and enlarge our QMC community.

\section{Sustainable} \label{CDHJS_sec:sustainable} 
Great QMC software should be alive.  It should be under active development or merged into a libraary that is actively maintained.  Commercial software has a infrastructure to sustain the algorithms in their libraries, but without attention, those algorithms can become stale.  Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.

@@

@@

}

@@

\section{Introduction} \label{CDHJS_sec:intro}
Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.  In some cases QMC  achieves one hundredth of the error of IID MC in the same amount of time. Often, these efficiency gains are obtained simply by  replacing IID sampling by low discrepancy (LD) sampling, which is the heart of QMC.

@@

\section{Available Software for QMC} \label{CDHJS_sec:available}
QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
Software focusing on generating high quality LD sequences  or their generators includes
\begin{description}[format=\textup,format=\textbf]
% https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
	\item[BRODA] Sobol' sequences in C, MATLAB, and Excel \cite{BRODA20a},
	\item[Burkhardt] various QMC software in C++, Fortran, MATLAB, \& Python \cite{Bur20a},
	\item[LatNet Builder] Generating vectors/matrices for lattices and digital nets \cite{LEcEtal22a,LatNet},
	\item[MATLAB] Sobol' and Halton sequences, commercial \cite{MAT9.9},
	\item[MPS] Magic Point Shop, lattices and Sobol' sequences \cite{Nuy17a},
	\item[Owen] Randomized Halton sequences in R \cite{Owe20a},
	\item[PyTorch] Scrambled Sobol' sequences \cite{PyTorch},
	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}, and
	\item [qrng]  Sobol', Halton, and Korobov sequences in R \cite{QRNG2020}.
\end{description}
Software focusing on QMC algorithms and applications includes
\begin{description}[format=\textup,format=\textbf]
	\item[GAIL] Automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal20a},
	\item[ML(Q)MC] Multi-Level (Quasi-)Monte Carlo routines in C++, MATLAB, Python, and R \cite{GilesSoft},
	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks 'N Statistics in Python \cite{OpenTURNS},
	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients \cite{KuoNuy16a},
	\item[SSJ] Stochastic Simulation in Java \cite{SSJ}, and
	\item[UQLab] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.
\end{description}

@@

The software libraries referenced in Section \ref{CDHJS_sec:available} provide one or more of these components. QMCPy combines multiple examples of them all under and object oriented framework. Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component. The following sections detail descriptions and use cases for each component and select implementations.

@@

\section{Stopping Criteria} \label{CDHJS_sec:stopping_crit}

@@

QMC, which uses LD sequences, is touted as providing substantially greater computational efficiency compared to IID MC.
Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences. Consistent with what is stated in Section \ref{CDHJS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.  By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.  We see that QMC methods often require orders of magnitude fewer samples that MC methods to achieve the same error tolerance.

@@

A helpful features of QMCPy is the ability to print out components instances to see more detailed information. This allows users to quickly see the properties that have been set and defaulted for any given component. This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHJS_sec:stopping_crit}. The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}

@@

\section{Further Work} \label{CDHJS_sec:further}

@@

\section{Tenets of Great Software} \label{CDHJS_sec:tenets_of_great_software}

Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication. A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement. Software accessibility considerations are further discussed in Sect.\ \ref{CDHJS_sec:language_library_environment}.

Libraries should be \emph{sustainable} for future development. The goal should be to have a sufficient user base and developer community for updates and maintenance. Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section. However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support. We promote collaborative, community driven development further in Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.

\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHJS_sec:qmc_software_architecture}

\section{Choosing a Environment} \label{CDHJS_sec:language_library_environment}

\section{Encouraging Collaborative Development} \label{CDHJS_sec:encouraging_collaborative_developement}

\section{Example Collaboration with UMBridge and QMCPy} \label{CDHJS_sec:example_collaboration_umbridge}

\section*{TODO}
\begin{itemize}
    \item Flesh out Section \ref{CDHJS_sec:encouraging_collaborative_developement}
    \item QMC Software Architecture section: better way then description of components and list of softwares?
    \item For Section \ref{CDHJS_sec:tenets_of_great_software}, the slides also include
    \begin{itemize}
        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
        \item Current---include the latest and best algorithms
    \end{itemize}
    which I think fall under the umbrella of software being \emph{integrated} with the community.
    \item update QMCPy citation \cite{QMCPy2020a}
    \item figures?
    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
\end{itemize}

Developing software to implement such enhanced MC strategies poses a number of challenges. First, a development team must determine the value of the software to the users. We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHJS_sec:why_we_need_software}. Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHJS_sec:tenets_of_great_software}. Third, an architecture for the  (Q)MC software should be determined. We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHJS_sec:qmc_software_architecture}. Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support. We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHJS_sec:language_library_environment}. Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community. To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.
